---
title: '<center> The Exercises of Week 4 - Classification and Model Evaluation <center>'
author: '<center> Reza Mohammadi <center>'
date: '<center> `r Sys.Date()` <center>'
output:
  html_document:
    number_sections: true
    fig_caption: true
    toc: true
    fig_width: 7
    fig_height: 5
    theme: cosmo
    highlight: tango
    code_folding: show
---

```{r setup, include = FALSE}
knitr::opts_chunk $ set( echo = TRUE, message = FALSE, warning = FALSE, 
                         comment = " ", error = FALSE, fig.align = 'center'  )
```


```{css, echo=FALSE}
.answer_panel { margin-left: 40px; }

.sub_details { margin-left: 60px; }
```

**Full Name:**

* Peter Molnar (11759216)
  
**Online Assignment**: There is a new online assignment at the [DataCamp](https://www.datacamp.com) with the name “*Chapter 1: k-Nearest Neighbors (kNN)*” which is a part of the online course “Supervised Learning in R: Classification” at the DataCamp. *The online assignments at the DataCamp are not mandatory*.

**Your task** is to answer the following questions in *Part 1* and *Part 2* in this R-markdown file. Please upload both your R-markdown (.Rmd file) and the HTML files separately on Canvas. Note that your R-markdown (.Rmd file) and the HTML files have to be in the right format.

Here, we are going to use the following R packages:

* **[liver](https://CRAN.R-project.org/package=liver)**: the *bank* and *churn* datasets are in this package. We also use the `partition()`, `kNN()` functions in this package.
* **[ggplot2](https://ggplot2.tidyverse.org/index.html)**: we use this package for visualization.
* **[pROC](https://CRAN.R-project.org/package=pROC)**: to create ROC curve and compute AUC values, we use the `roc()` and `auc()` functions in this package.

If it’s needed, install these packages on your computer. Here we load them:
```{r}
if( !require( liver   ) ) install.packages( "liver" )
if( !require( ggplot2 ) ) install.packages( "ggplot2" )
if( !require( pROC    ) ) install.packages( "pROC" )

library( liver )  
library( ggplot2 )   
library( pROC )
library( caret )
```

# Classification for Bank direct marketing dataset (40 points)

By using different classification algorithms, we want to detect customers who are more likely to acquire the product.

## Business Understanding

Find the best strategies to improve for the next marketing campaign. How can the financial institution have greater effectiveness for future marketing campaigns? To make a data-driven decision, we need to analyze the last marketing campaign the bank performed and identify the patterns that will help us find conclusions to develop future strategies.

### Bank direct marketing info

Two main approaches for enterprises to promote products/services are: 

* *mass campaigns*: targeting general indiscriminate public,
* *directed marketing*, targeting a specific set of contacts. 

In general, positive responses to mass campaigns are typically very low (less than 1%). On the other hand, direct marketing focuses on targets that are keener to that specific product/service, making this kind of campaign more effective. However, direct marketing has some drawbacks, for instance, it may trigger a negative attitude towards banks due to the intrusion of privacy.

Banks are interested to increase financial assets. One strategy is to offer attractive long-term deposit applications with good interest rates, in particular, by using directed marketing campaigns. Also, the same drivers are pressing for a reduction in costs and time. Thus, there is a need for an improvement in efficiency: lesser contacts should be done, but an approximate number of successes (clients subscribing to the deposit) should be kept.

### What is a Term Deposit?

A Term Deposit is a deposit that a bank or a financial institution offers with a fixed rate (often better than just opening a deposit account), in which your money will be returned at a specific maturity time. For more information with regards to Term Deposits please check [here](https://www.investopedia.com/terms/t/termdeposit.asp).

## Data Undestanding

The *bank* dataset is related to direct marketing campaigns of a Portuguese banking institution. You can find more information related to this dataset at: [https://rdrr.io/cran/liver/man/bank.html](https://rdrr.io/cran/liver/man/bank.html)

The marketing campaigns were based on phone calls. Often, more than one contact (to the same client) was required, to access if the product (bank term deposit) would be (or not) subscribed. The classification goal is to predict if the client will subscribe to a term deposit (variable deposit).

We import the *bank* dataset:
```{r}
data( bank )      
```

We can see the structure of the dataset by using the `str()` function:
```{r}
str( bank )
```

It shows that the *bank* dataset as a `data.frame` has `r ncol( bank )` variables and `r nrow( bank )` observations. The dataset has `r ncol( bank ) - 1` predictors along with the target variable `deposit` which is a binary variable with 2 levels "yes" and "no". The variables in this dataset are:

* `age`: numeric.
* `job`: type of job; categorical: "admin.", "unknown", "unemployed", "management", "housemaid", "entrepreneur", "student", "blue-collar, "self-employed", "retired", "technician", "services".
* `marital`: marital status; categorical: "married", "divorced", "single"; note: "divorced" means divorced or widowed.
* `education`: categorical: "secondary", "primary", "tertiary", "unknown".
* `default`: has credit in default?; binary: "yes","no".
* `balance`: average yearly balance, in euros; numeric.
* `housing`: has housing loan? binary: "yes", "no".
* `loan`: has personal loan? binary: "yes", "no".

Related with the last contact of the current campaign:

* `contact`: contact: contact communication type; categorical: "unknown","telephone","cellular". 
* `day`: last contact day of the month; numeric.
* `month`: last contact month of year; categorical: "jan", "feb", "mar", ..., "nov", "dec".
* `duration`: last contact duration, in seconds; numeric.

Other attributes:

* `campaign`: number of contacts performed during this campaign and for this client; numeric, includes last contact.
* `pdays`: number of days that passed by after the client was last contacted from a previous campaign; numeric, -1 means client was not previously contacted.
* `previous`: number of contacts performed before this campaign and for this client; numeric.
* `poutcome`: outcome of the previous marketing campaign; categorical: "success", "failure", "unknown", "other".

Target variable:

* `deposit`: Indicator of whether the client subscribed a term deposit; binary: "yes" or "no".

Here we report the summary of the dataset:
```{r}
summary( bank )
```

## Data Preparation

We partition the *bank* dataset randomly into two groups: train set (80%) and test set (20%). Here, we use the `partition()` function from the *liver* package:

```{r}
set.seed( 5 )

data_sets = partition( data = bank, prob = c( 0.8, 0.2 ) )

train_set = data_sets $ part1
test_set  = data_sets $ part2

actual_test  = test_set $ deposit
```

Note that here we are using the `set.seed()` function to create reproducible results. 

We want to validate the partition by testing whether the proportion of the target variable `deposit` differs between the two data sets. We use a Two-Sample Z-Test for the difference in proportions. To run the test, we use the `prop.test()` function in **R**:
```{r}
x1 = sum( train_set $ deposit == "yes" )
x2 = sum( test_set  $ deposit == "yes" )

n1 = nrow( train_set )
n2 = nrow( test_set  )

prop.test( x = c( x1, x2 ), n = c( n1, n2 ) )
```

Based on the output, answer the following questions:

a. **Why is the above hypothesis test suitable for the above research question? Provide your reasons.**

<details class="answer_panel">
  <summary>Answer a</summary>
Above we want to validate whether the two partitions, training set and test set are independent. Since the type of `deposit` is of binary/flag, and we aim to compare the two subsets, we carry out a two-sample Z-Test for the difference is proportions.
  
If the proportions are significantly different between training set and test set, then this may lead to bias or come at the cost of prediction/classification precision. Else, if the proportions are approximately the same, we can conclude that the training set is a representative set of the test set.

**Reasons:**

* 2 samples: training set, test set
* data type: binary/flag variables
* sufficiently large sample size $N_{train} = 3617, N_{test} = 904$, so that $Z \stackrel{approx}{\sim} N(\mu_{deposit}, \sigma_{depost})$

</details>

b. **Specify the null and alternative hypotheses?**

<details class="answer_panel">
  <summary>Answer b</summary>

\[
\bigg\{
\begin{matrix}
          H_0:  \pi_{training}  = \pi_{test} \\
          H_a:  \pi_{training}  \neq \pi_{test} \\
\end{matrix}
\]

</details>

c. **Explain that you reject or do not reject the null hypothesis, at $\alpha=0.05$. What would be your statistical conclusion?**

<details class="answer_panel">
  <summary>Answer c</summary>

**Statistical conclusion:** Given significance level of $\alpha = 0.05$, and $p-value = 0.1217 > \alpha = 0.05$, there is insufficient evidence found to reject $H_0$, and we infer that the population proportions of the number of people subscribed to the term deposit is not different between the training and the test partitions.

</details>

d. **What would be a non-statistical interpretation of your findings in c?**

<details class="answer_panel">
  <summary>Answer d</summary>

**Non-statistical conclusion:** At $\alpha = 0.05$, we infer that the population proportions of the number of people subscribed to the term deposit is not different between the training and the test partitions, and hence the training set is a representative partition of the data

</details>


## Classification using the kNN algorithm

The results from the "Exploratory Data Analysis (EDA)" (from last week) indicate that the following predictors from `r ncol( bank ) - 1` predictors in the *bank* dataset are important to predict `deposit`.

`age`, `default`, `balance`, `housing`, `loan`, `duration`, `campaign`, `pdays`, and `previous`.

Thus, here, based on the training dataset, we want to apply kNN algorithm, by using above predictors in our model. We use the following formula:
```{r}
formula = deposit ~ age + default + balance + housing + loan + duration + campaign + pdays + previous
```

**NOTE:** The above formula means `deposit` is the target variable and the rest of the variables in the right side of tilde ("`~`") are independent variables. 

Based on the training dataset, we want to find the k-nearest neighbor for the test data set. Here we use two different values for k (k = 3 and k = 10). We use the `kNN()` function from the *R* package **liver**:
```{r}
predict_knn_3  = kNN( formula, train = train_set, test = test_set, k = 3  )

predict_knn_10 = kNN( formula, train = train_set, test = test_set, k = 10 )
```

To have an overview of the prediction result, we report *Confusion Matrix* for two different values of k by using the `conf.mat` function: 
```{r}
( conf_knn_3 = conf.mat( predict_knn_3, actual_test ) )

( conf_knn_10 = conf.mat( predict_knn_10, actual_test ) )
```

We also could report *Confusion Matrix* by using the `conf.mat.plot()` command:
```{r fig.show = "hold", out.width = "50%", fig.align = 'default'}
conf.mat.plot( predict_knn_3, actual_test, main = "kNN with k = 3" )

conf.mat.plot( predict_knn_10, actual_test, main = "kNN with k = 10" )
```

**What do these values mean? Explain what conclusion you will draw.**

<details>
  <summary>Answer</summary>
  
* There are 2 possible predicted classes: "yes" and "no"
  * "yes" meaning that we predicted that the targeted person would subscribe to the term deposit
  * "no" meaning that we predicted that the targeted person would not subscribe to the term deposit
* The classifier made a total of 946 predictions

<h3>k=3</h3>

The classifier 

* predicted "no" for 882 people
* predicted "yes" for 64 people
* correctly predicted the outcome 806 + 19 = 825 times, these values are true positives
* incorrectly predicted the outcome 45 + 76 = 121 times, these values are false negatives
  
Below we report a list of model evaluation ratios to draw a conclusion:
```{r}
confusionMatrix(data = predict_knn_3, reference = actual_test)
```
We identify the following:

* **Accuracy**: overall the classifier is correct 87.21% of the time
* **Misclassification rate**: overall the classifier is wrong 12.79% of the time
* **Sensitivity**: The classifier predicts "yes"-es 94.71% of the time
* **Specificity**: The classifier correctly predicts "no"-s 20.00% of the time

<h3>k=10</h3>
The classifier 

* predicted "no" for 908 people
* predicted "yes" for 38 people
* correctly predicted the outcome 834 + 21 = 855 times, these values are true positives
* incorrectly predicted the outcome 17 + 74 = 91 times, these values are false negatives
  
Below we report a list of model evaluation ratios to draw a conclusion:
```{r}
confusionMatrix(data = predict_knn_10, reference = actual_test)
```
We identify the following:

* **Accuracy**: overall the classifier is correct 90.38% of the time
* **Misclassification rate**: overall the classifier is wrong 0.0962% of the time
* **Sensitivity**: The classifier predicts "yes"-es 98.00% of the time
* **Specificity**: The classifier correctly predicts "no"-s 22.11% of the time
</details>


<h3>Conclusion</h3>

Based on the rate of accuracy, sensitivity, and specificity, the kNN algorithm with k=10 is a better predictor.

## Model evaluation by MSE

To evaluate the accuracy of the predictions, we calculate the Mean Square Error (MSE) by using the `mse()` function from the **liver** package:

```{r}
MSE_3 = mse( predict_knn_3, actual_test  )
MSE_3 

MSE_10 = mse( predict_knn_10, actual_test )
MSE_10
```

**For the case k=3, the MSE = `r round( MSE_3 , 3 )` and for the case k = 10, the MSE = `r round( MSE_10, 3 )`. What do these values mean? Explain what conclusion you will draw.**

<details>
  <summary>Answer</summary>

The Mean Squared Error combines the variance and bias into one formula and measures the squared difference between the observed value and the predicted value of a continuous target variable. Given that in our case that target variable is binary, the interpretation may be inaccurate.

For k=3, MSE of 0.128 means that the on average our prediction is approximately 12.8% off.
For k=10, MSE of 0.096 means that the on average our prediction is approximately 9.6% off.

Between the two competing models, the one with the lower MSE is preferred, that is kNN with k=10.
  
</details>

## Visualizing Model Performance by ROC curve

To report the ROC curve, we need the probability of our classification prediction. We can have it by using:
```{r}
prob_knn_3  = kNN( formula, train = train_set, test = test_set, k = 3 , type = "prob" )[ , 1 ]

prob_knn_10 = kNN( formula, train = train_set, test = test_set, k = 10, type = "prob" )[ , 1 ]
```

To visualize the model performance, we could report the ROC curve plot by using the `plot.roc()` function from the **pROC** package:

```{r}
roc_knn_3 = roc( actual_test, prob_knn_3 )

roc_knn_10 = roc( actual_test, prob_knn_10 )

ggroc( list( roc_knn_3, roc_knn_10 ), size = 0.8 ) + 
    theme_minimal() + ggtitle( "ROC plots with AUC for kNN") +
    scale_color_manual( values = c( "red", "blue" ), 
    labels = c( paste( "k=3 ; AUC=", round( auc( roc_knn_3  ), 3 ) ),
                paste( "k=10; AUC=", round( auc( roc_knn_10 ), 3 ) )
                ) ) +
    theme( legend.title = element_blank() ) +
    theme( legend.position = c( .7, .3 ), text = element_text( size = 17 ) )
```


In the above plot, '<span style="color:red">red</span>' curve is for the case k = 3 and the '<span style="color:blue">blue</span>' curve is for the case k = 10.

**Explain what conclusion you will draw. Do we need to report AUC (Area Under the Curve) as well?**

<details class="answer_panel">
  <summary>Answer</summary>
  
In question 1.4, we reported that kNN with k=10 has higher sensitivity and specificity ratios than kNN with k=3. This means, that the former curve will always lie above (except in sensitivity=1 and specificity=1) the latter, and hence will have a larger area enclosed under the curve.

The conclusion is the same as above, given that the area under the blue curve is greater than for red, we infer that the blue k=10 bears with smaller bias and smaller variance, and hence a better predictor than kNN with k=3.

Do we need to report AUC???

</details>

## kNN algorithm with data transformation

The predictors that we used in the previous question, do not have the same scale. For example, variable `duration` change between `r min( bank $ duration )` and `r max( bank $ duration )`, whereas the variable `loan` is binary. In this case, the values of variable `duration` will overwhelm the contribution of the variable `loan`. To avoid this situation we use normalization. So, we use min-max normalization and transfer the predictors. 

Now, we find the k-nearest neighbor for the test set, based on the training dataset, for the k = 10:

```{r}
predict_knn_10_trans = kNN( formula, train = train_set, test = test_set, transform = "minmax", k = 10 )

conf.mat.plot( predict_knn_10_trans, actual_test )
```

## ROC curve and AUC for transformed data

To report the ROC curve, we need the probability of our classification prediction. We can have it by using:
```{r}
prob_knn_10 = kNN( formula, train = train_set, test = test_set, k = 10, type = "prob" )[ , 1 ]

prob_knn_10_trans = kNN( formula, train = train_set, test = test_set, transform = "minmax", k = 10, type = "prob" )[ , 1 ]
```

To visualize the model performance between the raw data and the transformed data, we could report the ROC curve plot as well as AUC (Area Under the Curve) by using the `plot.roc()` function from the **pROC** package:

```{r}
roc_knn_10 = roc( actual_test, prob_knn_10 )

roc_knn_10_trans = roc( actual_test, prob_knn_10_trans )

ggroc( list( roc_knn_10, roc_knn_10_trans ), size = 0.8 ) + 
    theme_minimal() + ggtitle( "ROC plots with AUC for kNN") +
    scale_color_manual( values = c( "red", "blue" ), 
      labels = c( paste( "Raw data             ; AUC=", round( auc( roc_knn_10 ), 3 ) ), 
                  paste( "Transformed data; AUC=", round( auc( roc_knn_10_trans ), 3 ) ) ) ) +
  theme( legend.title = element_blank() ) +
  theme( legend.position = c( .7, .3 ), text = element_text( size = 17 ) )
```

In the above plot black curve is for the *raw* dataset and the red curve is for the *transformed* dataset.

**Explain what conclusion you will draw. Based on the values of AUC (Area Under the Curve), explain what conclusion you will draw.**

<details class="answer_panel">
  <summary>Answer</summary>
  
Given that the different variables are now standardised, none of them will overwhelm the contribution of others. We observe, that AUC for k=10 has improved by 0.017, result in in an AUC of 0.814 for the transformed data.

Between competing models, the one with higher AUC is preferred, therefore, the transformed data is deemed a better predictor.

</details>

## Optimal value of k for the kNN algorithm

In the previous questions, for finding the k-nearest neighbor for the test set, we set k = 10. But why 10? Here, we want to find out the optimal value of k based on our dataset. 

To find out the optimal value of `k` based on *Error Rate*, for the different values of k from 1 to 30, we run the k-nearest neighbor for the test set and compute the *Error Rate* for these models, by running `kNN.plot()` command 

```{r}
kNN.plot( formula, train = train_set, test = test_set, transform = "minmax", 
          k.max = 30, set.seed = 7 )
```

**Based on the plot, what value of k is optimal? Provide your reasons.**

<details class="answer_panel">
  <summary>Answer</summary>
  
Between competing models, the one with the lowest error rate is preferred over others. That is, the optimal value of k for the `bank` dataset is k=8.

</details>

# Applying kNN algorithm for *churn* dataset (60 points)

Apply the kNN algorithm to analyze the *churn* dataset which is available in the **R** package [**liver**](https://CRAN.R-project.org/package=liver). 

The R code for this part is available in the file "Rcode_W4_part2.R" at "Canvas > Modules > Week 4 - Classification \& Model Evaluation".
   
## Import the *churn* dataset

Import the *churn* dataset in R and report the summary of the dataset. 

## Data Preparation

First, partition the *churn* dataset randomly into two groups as a train set (80%) and test set (20%). Then, validate the partition for a couple of variables; for example, you could validate the partition by testing whether the proportion of churners differs between the two datasets. Or you could validate the partition by testing whether the population means for the number of customer service calls differs between the two datasets.

## Applying the kNN algorithm using all predictors

Based on the training dataset, find the k-nearest neighbor for the test data set, for the case k = 10; For this question, use all the 19 predictors of the *churn* dataset for the analysis. Note that you should use min-max normalization and transfer the predictors. Then, evaluate the accuracy of the predictions by reporting 

* Confusion Matrix
* MSE
* ROC curve
* AUC

## Applying the KNN algorithm using part of predictors

In the previous exercises for the *churn* dataset, you suppose to use all the 19 predictors for the analysis. But we know based on the lecture of week 2, we should use only those predictors that have a relationship with the target variable. So, here we use the following predictors:

`account.length`, `voice.plan`, `voice.messages`, `intl.plan`, `intl.mins`, `day.mins`, `eve.mins`, `night.mins`, and `customer.calls`.

First, based on the above predictors, find the k-nearest neighbor for the test set, based on the training dataset, for the k = 10. Note that you should use min-max normalization and transfer the predictors. Then, evaluate the accuracy of the predictions by reporting 

* Confusion Matrix
* MSE
* ROC curve
* AUC

Compare the results with the previous section. What would be your conclusion? 

## Optimal value of k for the kNN algorithm

In the previous questions, to find the k-nearest neighbor for the test set, we set k = 10. But why 10? Find out the optimal value of `k`.
   
# **Bonus**: Applying the kNN algorithm for your own dataset (30 points)

In this part, apply the kNN algorithm to analyze your own dataset. You could follow the same steps as in part 1 (above) of these exercises.

Note: to apply the kNN algorithm to your own dataset, the target variable has to be categorical, preferably Binary, similar to the above example.
  